---
title: "Introduction to the tidytopic package"
author: "Mans Magnusson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tidytopics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The tidytopics package is a toolkit for topic modeling and analysis in R. This is a short introduction to the package, the data structures and some basic functionalities. See 

## Basic data structure

The datastructures in tidytopics is based on the same tidy data structure proposed by Hadley Wickham in tidyr. There are two main data structures in tidytopics: `tidy_topic_state` and `tidy_topic_matrix`/`tidy_topic_array`. This is a way of handling the often very large corpuses in a simple and effective way for post processing and analysis using the `tidytext` and `dplyr` R packages.

EXAMPLES HERE sparse matrix 


### Example: A tidy type probability matrix

Examples of type_probabilities

Examples of topic_probabilities


## Import and export mallet state files

```{r}
library(tidytopics)
library(tidytext)
library(janeaustenr)
library(dplyr)
library(stringr)

original_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]",
                                                 ignore_case = TRUE)))) %>%
  ungroup()

tidy_books <- original_books %>%
  unnest_tokens(word, text)

```


## Session info

This vignette was created with

```{r sessioninfo, message=FALSE, warning=FALSE}
sessionInfo()
```
